Effects of diversity, topography, and interannual climate variability on pathogen spillover

### 2016-02-25 esa-2016-abstract.Rmd
Update backgroun/methods.

## Background/Question/Methods
Variation in disease transmission is caused by heterogeneity of individuals, species, and landscapes across space and time. Multi-host disease systems are often characterized by pathogen spillover due to differences between host-pathogen interactions among the various hosts. Sudden oak death is a multi-host forest disease characterized by pathogen spillover, where a single host species (California bay laurel) amplifies the pathogen, which spills over to infect nearby oak trees. We applied a path analysis to analyze the direct and indirect effects of diversity, topography, and interannual climate variability on pathogen spillover for sudden oak death. The path analysis is a set of hypothesis built on findings from experiments and observations of the disease. We applied the path analysis framework to 10 years of data from a network of disease-monitoring plots distributed across a 275-km2 area in southeastern Sonoma County, California. During each year, we measured temperature, rainfall, pathogen load, and oak infection, while plant community diversity was measured once at each plot in this low-turnover landscape.

>>>>

## Background/Question/Methods
Variation in disease transmission is caused by heterogeneity of individuals, species, and landscapes across space and time. For example, multi-host disease systems are often characterized by pathogen spillover due to differences between host-pathogen interactions among the various hosts. Understanding the complexities of these disease dynamics at meaningful scales is a continuing challenge for disease ecologists. Sudden oak death is a multi-host forest disease characterized by pathogen spillover, where a single host species (California bay laurel) amplifies the pathogen, which spills over to infect nearby oak trees. We used path analysis to analyze the direct and indirect effects of diversity, topography, and interannual climate variability on pathogen spillover (oak infection). The hypothesized relationships laid out by our path model for sudden oak death were based on published findings from experiments and observations of the disease during the past 15 years. We applied our path analysis framework to 10 years of data from a network of disease-monitoring plots distributed across a 275-km2 area in southeastern Sonoma County, California. During each year, we measured temperature, rainfall, pathogen load, and oak infection, while plant community diversity was measured once at each plot in this low-turnover landscape.
